---
title: "PhilTrans_MaternalUncles"
author: "MHKeith"
date: "October 12, 2018"
output: html_document
---

Load packages, data, and run summary statistics for Table 1. 

```{r setupstats}

#load file "ShodagorMatUncleDat.csv" from working directory
MUDat<-read.csv("ShodagorMatUncleDat.csv")

#summary stats for numerical variables, Table 1A
stats1A<-apply(MUDat[ ,c(2,4,7,18,20,22,25,27,29,31,33,35)],2,summary)
Tab1A<-t(stats1A)
Tab1ASD<-data.frame(apply(MUDat[ ,c(2,4,7,18,20,22,25,27,29,31,33,35)],2,sd))
colnames(Tab1ASD)<-"SD"
Table1A<-cbind(Tab1A,Tab1ASD)

#summary stats for dummy variables, Table 1B
Tab1B<-apply(MUDat[ ,c(6,39,40,9,10,11,12,13,14,16,15,17,24)],2,table)
Table1B<-t(Tab1B)
colnames(Table1B)<-c("No", "Yes")

```

Screen predictors in clusters, and retain variables with 90% credible intervals that do not span zero for final models. 

```{r predscreen}

library(doParallel)
library(MCMCglmm)

#clusters A-E, screen separately

cl <- makeCluster(3) #3 models, run them in parallel on 3 cores
registerDoParallel(cl)

#use fixed prior on residuals (unidientifiable in logit models)- list R
#use regularizing prior on fixed effects to prevent overfitting- list B

priorA <- list(R = list(V = 1, fix=1),
               B = list(mu = c(rep(0,8)), V = diag(8)*(4.3+pi^2/3)))

#Cluster A
system.time(modsA <- foreach(i=1:3) %dopar% {
  library(MCMCglmm)
    forml<-list(as.formula(MU_InvestAny ~ Z_MOAge +IncomeLog +Boat_dum + MatRes + PatRes + OthRes + Z_AnnMove + 1), 
                as.formula(MU_Direct ~ Z_MOAge +IncomeLog+Boat_dum + MatRes + PatRes + OthRes + Z_AnnMove  + 1), 
                as.formula(MU_Indirect_dum ~ Z_MOAge +IncomeLog+Boat_dum + MatRes + PatRes + OthRes + Z_AnnMove + 1))

    MCMCglmm( forml[[i]],
        data=MUDat,
        prior=priorA,
        family='categorical',
        nitt = 1100000, thin = 1000, burnin = 100000,
        pl=TRUE, pr=TRUE,
        verbose=FALSE)
})


# function for marginal R2 from MCMCglmm models
mcmcR2 <- function(m){
    X <- as.matrix(m$X)
    sigF <- apply(m$Sol[,1:ncol(X)], 1, function(z) var(as.vector(z %*% t(X))) )
    R2m <- sigF / ( sigF + rowSums(m$VCV) )
    if(ncol(m$VCV)==2){nv<-m$VCV[,-which(colnames(m$VCV)=="units")]}
    else{nv<-rowSums(m$VCV[,-which(colnames(m$VCV)=="units")])}
    outmc <- matrix(NA,nrow=length(sigF),ncol=2)
    outmc[,1] <- sigF
    outmc[,2] <- R2m
    colnames(outmc) <- c("sigF","R2m")
    class(outmc) <- "mcmc"
    attr(outmc,"mcpar") <- attr(m$Sol,"mcpar")
    return( outmc )
}

AnyA<-modsA[[1]]
summary(AnyA)
anyAresults<-data.frame(posterior.mode(AnyA$Sol), HPDinterval(AnyA$Sol,prob=0.90))
#retain boat dummy

DirA<-modsA[[2]]
summary(DirA)
dirAresults<-data.frame(posterior.mode(DirA$Sol), HPDinterval(DirA$Sol,prob=0.90))
#retain number of annual movements, boat dummy 

IndA<-modsA[[3]]
summary(IndA)
indAresults<-data.frame(posterior.mode(IndA$Sol), HPDinterval(IndA$Sol,prob=0.90))

#R2 A cluster
R2A<-as.data.frame(lapply(modsA,mcmcR2)) 
names(R2A)<-c("sigFany", "R2many","sigFdir","R2mdir","sigFind","R2mind") 
anyR2Amode<-posterior.mode(R2A$R2many) #marginal R2 posterior mode for "any" model
anyR2ACI<-quantile(R2A$R2many,c(.10,.90)) #marginal R2 90% credible interval for "any" model
dirR2Amode<-posterior.mode(R2A$R2mdir) #direct mode
dirR2ACI<-quantile(R2A$R2mdir,c(.10,.90)) #direct CI
indR2Amode<-posterior.mode(R2A$R2mind) #indirect mode
indR2ACI<-quantile(R2A$R2mind,c(.10,.90)) #indirect CI

#Cluster B
priorB<- list(R = list(V = 1, fix=1),
               B = list(mu = c(rep(0,5)), V = diag(5)*(4.3+pi^2/3)))
  
system.time(modsB <- foreach(i=1:3) %dopar% {
  library(MCMCglmm)
    forml<-list(as.formula(MU_InvestAny ~ Z_MOAge + IncomeLog+ MOEdu_dum + IncomeLog + MOSells_Occup + 1), 
                as.formula(MU_Direct ~ Z_MOAge + IncomeLog+MOEdu_dum + IncomeLog + MOSells_Occup + 1), 
                as.formula(MU_Indirect_dum ~ Z_MOAge + IncomeLog+MOEdu_dum + IncomeLog + MOSells_Occup + 1))

    MCMCglmm( forml[[i]],
        data=MUDat,
        prior=priorB,
        family='categorical',
        nitt = 1100000, thin = 1000, burnin = 100000,
        pl=TRUE, pr=TRUE,
        verbose=FALSE)
})

AnyB<-modsB[[1]]
summary(AnyB)
anyBresults<-data.frame(posterior.mode(AnyB$Sol), HPDinterval(AnyB$Sol,prob=0.90))
#retain mother's occupation

DirB<-modsB[[2]]
summary(DirB)
dirBresults<-data.frame(posterior.mode(DirB$Sol), HPDinterval(DirB$Sol,prob=0.90))
#retain mother's occupation

IndB<-modsB[[3]]
summary(IndB)
indBresults<-data.frame(posterior.mode(IndB$Sol), HPDinterval(IndB$Sol,prob=0.90))

#R2 B cluster
R2B<-as.data.frame(lapply(modsB, mcmcR2))
names(R2B)<-c("sigFany", "R2many","sigFdir","R2mdir","sigFind","R2mind") 
anyR2Bmode<-posterior.mode(R2B$R2many) #any mode
anyR2BCI<-quantile(R2B$R2many,c(.10,.90)) #any CI
dirR2Bmode<-posterior.mode(R2B$R2mdir) #direct mode
dirR2BCI<-quantile(R2B$R2mdir,c(.10,.90)) #direct CI
indR2Bmode<-posterior.mode(R2B$R2mind) #indirect mode
indR2BCI<-quantile(R2B$R2mind,c(.10,.90)) #indirect CI

#Cluster C
priorC<- list(R = list(V = 1, fix=1),
               B = list(mu = c(rep(0,5)), V = diag(5)*(4.3+pi^2/3)))

system.time(modsC <- foreach(i=1:3) %dopar% {
    library(MCMCglmm)
    forml<-list(as.formula(MU_InvestAny ~ Z_MOAge + IncomeLog+ MONum_Marriages + FatherPresent + 1), 
                as.formula(MU_Direct ~ Z_MOAge + IncomeLog+ MONum_Marriages + FatherPresent + 1), 
                as.formula(MU_Indirect_dum ~ Z_MOAge + IncomeLog+ MONum_Marriages + FatherPresent + 1))

    MCMCglmm( forml[[i]],
        data=MUDat,
        prior=priorC,
        family='categorical',
        nitt = 1100000, thin = 1000, burnin = 100000,
        pl=TRUE, pr=TRUE,
        verbose=FALSE)
})

AnyC<-modsC[[1]]
summary(AnyC)
anyCresults<-data.frame(posterior.mode(AnyC$Sol), HPDinterval(AnyC$Sol,prob=0.90))

DirC<-modsC[[2]]
summary(DirC)
dirCresults<-data.frame(posterior.mode(DirC$Sol), HPDinterval(DirC$Sol,prob=0.90))

IndC<-modsC[[3]]
summary(IndC)
indCresults<-data.frame(posterior.mode(IndC$Sol), HPDinterval(IndC$Sol,prob=0.90))

#R2 C cluster
R2C<-as.data.frame(lapply(modsC, mcmcR2))
names(R2C)<-c("sigFany", "R2many","sigFdir","R2mdir","sigFind","R2mind") 
anyR2Cmode<-posterior.mode(R2C$R2many) #any mode
anyR2CCI<-quantile(R2C$R2many,c(.10,.90)) #any CI
dirR2Cmode<-posterior.mode(R2C$R2mdir) #direct mode
dirR2CCI<-quantile(R2C$R2mdir,c(.10,.90)) #direct CI
indR2Cmode<-posterior.mode(R2C$R2mind) #indirect mode
indR2CCI<-quantile(R2C$R2mind,c(.10,.90)) #indirect CI


#Cluster D
priorD <- list(R = list(V = 1, fix=1),
               B = list(mu = c(rep(0,7)), V = diag(7)*(4.3+pi^2/3))) 

system.time(modsD <- foreach(i=1:3) %dopar% {
  library(MCMCglmm)
    forml<-list(as.formula(MU_InvestAny ~ Z_MOAge + IncomeLog+ Z_MO_NumCldLiv + Z_MO_AgeFirstBirth + MO_LostChild_dum + Z_IBI_AfterFirst + 1), 
                as.formula(MU_Direct ~ Z_MOAge + IncomeLog+ Z_MO_NumCldLiv + Z_MO_AgeFirstBirth + MO_LostChild_dum + Z_IBI_AfterFirst + 1), 
                as.formula(MU_Indirect_dum ~ Z_MOAge + IncomeLog+ Z_MO_NumCldLiv + Z_MO_AgeFirstBirth + MO_LostChild_dum + Z_IBI_AfterFirst + 1))

    MCMCglmm( forml[[i]],
        data=MUDat,
        prior=priorD,
        family='categorical',
        nitt = 1100000, thin = 1000, burnin = 100000,
        pl=TRUE, pr=TRUE,
        verbose=FALSE)
})

AnyD<-modsD[[1]]
summary(AnyD)
anyDresults<-data.frame(posterior.mode(AnyD$Sol), HPDinterval(AnyD$Sol,prob=0.90))
#retain IBI

DirD<-modsD[[2]]
summary(DirD)
dirDresults<-data.frame(posterior.mode(DirD$Sol), HPDinterval(DirD$Sol,prob=0.90))

IndD<-modsD[[3]]
summary(IndD)
indDresults<-data.frame(posterior.mode(IndD$Sol), HPDinterval(IndD$Sol,prob=0.90))
#retain IBI, mother has lost a child

#R2 D cluster
R2D<-as.data.frame(lapply(modsD, mcmcR2))
names(R2D)<-c("sigFany", "R2many","sigFdir","R2mdir","sigFind","R2mind") 
anyR2Dmode<-posterior.mode(R2D$R2many) #any mode
anyR2DCI<-quantile(R2D$R2many,c(.10,.90)) #any CI
dirR2Dmode<-posterior.mode(R2D$R2mdir) #direct mode
dirR2DCI<-quantile(R2D$R2mdir,c(.10,.90)) #direct CI
indR2Dmode<-posterior.mode(R2D$R2mind) #indirect mode
indR2DCI<-quantile(R2D$R2mind,c(.10,.90)) #indirect CI

#Cluster E
priorE <- list(R = list(V = 1, fix=1),  
               B = list(mu = c(rep(0,8)), V = diag(8)*(4.3+pi^2/3))) 

system.time(modsE <- foreach(i=1:3) %dopar% {

    forml<-list(as.formula(MU_InvestAny ~ Z_MOAge + IncomeLog+ MO_Firstborn_dum + Z_MO_NumSibs + Z_MO_NumSisters + Z_MO_NumBro_Total + Z_MO_Num_OlderBro + Z_MO_Num_YoungBro + Z_Num_Alloparents + 1), 
                as.formula(MU_Direct ~ Z_MOAge + IncomeLog+ MO_Firstborn_dum + Z_MO_NumSibs + Z_MO_NumSisters + Z_MO_NumBro_Total + Z_MO_Num_OlderBro + Z_MO_Num_YoungBro + Z_Num_Alloparents + 1), 
                as.formula(MU_Indirect_dum ~ Z_MOAge + IncomeLog+ MO_Firstborn_dum + Z_MO_NumSibs + Z_MO_NumSisters + Z_MO_NumBro_Total + Z_MO_Num_OlderBro + Z_MO_Num_YoungBro + Z_Num_Alloparents + 1))

    MCMCglmm( forml[[i]],
        data=MUDat,
        prior=priorE,
        family='categorical',
        nitt = 1100000, thin = 1000, burnin = 100000,
        pl=TRUE, pr=TRUE,
        verbose=FALSE)
})

AnyE<-modsE[[1]]
summary(AnyE)
anyEresults<-data.frame(posterior.mode(AnyE$Sol), HPDinterval(AnyE$Sol,prob=0.90))
#retain mother firstborn, number of older brothers

DirE<-modsE[[2]]
summary(DirE)
dirEresults<-data.frame(posterior.mode(DirE$Sol), HPDinterval(DirE$Sol,prob=0.90))
#retain number of older brothers

IndE<-modsE[[3]]
summary(IndE)
indEresults<-data.frame(posterior.mode(IndE$Sol), HPDinterval(IndE$Sol,prob=0.90))
#retain mother firstborn 

#R2 E cluster
R2E<-as.data.frame(lapply(modsE, mcmcR2))
names(R2E)<-c("sigFany", "R2many","sigFdir","R2mdir","sigFind","R2mind") 
anyR2Emode<-posterior.mode(R2E$R2many) #any mode
anyR2ECI<-quantile(R2E$R2many,c(.10,.90)) #any CI
dirR2Emode<-posterior.mode(R2E$R2mdir) #direct mode
dirR2ECI<-quantile(R2E$R2mdir,c(.10,.90)) #direct CI
indR2Emode<-posterior.mode(R2E$R2mind) #indirect mode
indR2ECI<-quantile(R2E$R2mind,c(.10,.90)) #indirect CI

```

Final GLMs for outcomes any investment, direct investment, and indirect investment. 

```{r GLMs}

#Any investment
priorAny <- list(R = list(V = 1, fix=1),
              B = list(mu = c(rep(0,8)), V = diag(8)*(4.3+pi^2/3)))

system.time(Any <- {
  library(MCMCglmm)

    forml<-as.formula(MU_InvestAny ~ Z_MOAge + IncomeLog + MOSells_Occup + Boat_dum+ Z_IBI_AfterFirst + MO_Firstborn_dum + Z_MO_Num_OlderBro + 1)
    
    MCMCglmm( forml,
        data=MUDat,
        prior=priorAny,
        family='categorical',
        nitt = 10500000, thin = 10000, burnin = 500000,
        pl=TRUE, pr=TRUE,
        verbose=FALSE)
})

summary(Any)
anytab<-data.frame(posterior.mode(Any$Sol), HPDinterval(Any$Sol,prob=0.90))
AnyOR<-exp(Any$Sol) #transform to odds ratio scale for supplemental results
TableSIany<-data.frame(posterior.mode(AnyOR), HPDinterval(AnyOR,prob=0.90))
TableSIany$type<-rep("Any",8)
TableSIany$variable<-rownames(TableSIany)
names(TableSIany)[1]<-"mode"

#R2 Any final model
R2any<-as.data.frame(mcmcR2(Any))
names(R2any)<-c("sigFany","R2many")
anyR2fmode<-posterior.mode(R2any$R2many)
anyR2fCI<-quantile(R2any$R2many,c(0.10,0.90))

#Direct investment
priorDir <- list(R = list(V = 1, fix=1),
              B = list(mu = c(rep(0,7)), V = diag(7)*(4.3+pi^2/3)))

system.time(Direct <- {
  library(MCMCglmm)
    forml<-as.formula(MU_Direct ~ Z_MOAge + IncomeLog + MOSells_Occup + Boat_dum + Z_AnnMove + Z_MO_Num_OlderBro + 1)

    MCMCglmm( forml,
        data=MUDat,
        prior=priorDir,
        family='categorical',
        nitt = 10500000, thin = 10000, burnin = 500000,
        pl=TRUE, pr=TRUE,
        verbose=FALSE)
})

summary(Direct)
dirtab<-data.frame(posterior.mode(Direct$Sol), HPDinterval(Direct$Sol,prob=0.90))
DirectOR<-exp(Direct$Sol) #transform to odds ratio scale for Table S4
TableSIdirect<-data.frame(posterior.mode(DirectOR), HPDinterval(DirectOR,prob=0.90))
TableSIdirect$type<-rep("Direct",7)
TableSIdirect$variable<-rownames(TableSIdirect)
names(TableSIdirect)[1]<-"mode"

#R2 Direct final model
R2direct<-as.data.frame(mcmcR2(Direct))
names(R2direct)<-c("sigFdir","R2mdir")
dirR2fmode<-posterior.mode(R2direct$R2mdir)
dirR2fCI<-quantile(R2direct$R2mdir,c(0.10,0.90))

#Indirect investment
priorInd <- list(R = list(V = 1, fix=1),
              B = list(mu = c(rep(0,6)), V = diag(6)*(4.3+pi^2/3)))

system.time(Indirect <- {
  library(MCMCglmm)
    forml<-as.formula(MU_Indirect_dum ~Z_MOAge + IncomeLog+ MO_LostChild_dum + Z_IBI_AfterFirst + MO_Firstborn_dum + 1)

    MCMCglmm( forml,
        data=MUDat,
        prior=priorInd,
        family='categorical',
        nitt = 10500000, thin = 10000, burnin = 500000,
        pl=TRUE, pr=TRUE,
        verbose=FALSE)
})


summary(Indirect)
indtab<-data.frame(posterior.mode(Indirect$Sol), HPDinterval(Indirect$Sol,prob=0.90))
IndirectOR<-exp(Indirect$Sol) #transform to odds ratio scale for Table S4
TableSIindirect<-data.frame(posterior.mode(IndirectOR), HPDinterval(IndirectOR,prob=0.90))
TableSIindirect$type<-rep("Indirect",6)
TableSIindirect$variable<-rownames(TableSIindirect)
names(TableSIindirect)[1]<-"mode"

#R2 Indirect final model
R2indirect<-as.data.frame(mcmcR2(Indirect))
names(R2indirect)<-c("sigFind","R2mind")
indR2fmode<-posterior.mode(R2indirect$R2mind)
indR2fCI<-quantile(R2indirect$R2mind,c(0.10,0.90))

#Table S4
TableSI<-rbind(TableSIany,TableSIdirect,TableSIindirect)

```

Coefficient dot plots show Bayesian posterior modes and 90% credible intervals.Figures 1-3

```{r dotplots}

library(ggplot2)

#Figure 1
rownames(dirtab)<-c("Intercept","Mother's age (z)", "Household income (log)", "Mother sells goods", "Household lives on a boat","Number of annual movements (z)"  ,"Mother's no. of older brothers (z)")
dirtab$variable<-rownames(dirtab)
dirplot<-dirtab[-c(1), ]
names(dirplot)[1]<-"coefficient"
dirdotplot<-ggplot(data = dirplot, aes(x = variable, y = coefficient, ymin = lower, ymax = upper)) 
dirdotplot +  geom_point(size=2.5,position = position_dodge(width = 0.65)) +
  geom_errorbar(position = position_dodge(width = 0.65), width = 0.4)  + ylim(-3.5,4.5) + ylab("Coefficient") + xlab(NULL) + ggtitle("Log Odds of Direct MU Investment") +theme_minimal() + geom_hline(yintercept=0) + coord_flip()

#Figure 2
rownames(indtab)<-c("Intercept","Mother's age (z)", "Household income (log)", "Mother has lost a child", "IBI after first birth (z)", "Mother was firstborn")
indtab$variable<-rownames(indtab)
indplot<-indtab[-c(1), ]
names(indplot)[1]<-"coefficient"
inddotplot<-ggplot(data = indplot, aes(x = variable, y = coefficient, ymin = lower, ymax = upper)) 
inddotplot +  geom_point(size=2.5,position = position_dodge(width = 0.65)) +
  geom_errorbar(position = position_dodge(width = 0.65), width = 0.4)  + ylab("Coefficient") + xlab(NULL) + ggtitle("Log Odds of Indirect MU Investment") +theme_minimal() + geom_hline(yintercept=0) + coord_flip()

#Figure 3
rownames(anytab)<-c("Intercept","Mother's age (z)", "Household income (log)", "Mother sells goods", "Household lives on a boat", "IBI after first birth (z)", "Mother was firstborn", "Mother's no. of older brothers (Z)")
anytab$variable<-rownames(anytab)
anyplot<-anytab[-c(1), ]
names(anyplot)[1]<-"coefficient"
anydotplot<-ggplot(data = anyplot, aes(x = variable, y = coefficient, ymin = lower, ymax = upper)) 
anydotplot +  geom_point(size=2.5,position = position_dodge(width = 0.65)) +
  geom_errorbar(position = position_dodge(width = 0.65), width = 0.4)  + ylim(-3.0,7.0) + ylab("Coefficient") + xlab(NULL) + ggtitle("Log Odds of Any MU Investment") +theme_minimal() + geom_hline(yintercept=0) + coord_flip()

```

Statistics for maternal uncles providing investment 

```{r matunclestats}

library(vcd)

#subset to investing maternal uncles
MUsub<-subset(MUDat, MU_InvestAny==1)

#summary stats for numeric variables, Table 2
stats2A<-apply(MUsub[ ,37:38],2,summary)
Tab2A<-t(stats2A)
Tab2ASD<-data.frame(apply(MUsub[ ,37:38],2,sd))
colnames(Tab2ASD)<-"SD"
Table2A<-cbind(Tab2A,Tab2ASD)

#summary stats for dummy variables, Table 2
Table2B<-apply(MUsub[ ,39:52],2,table)
Tab2B<-t(Table2B)
colnames(Tab2B)<-c("No", "Yes")


#mosaic plots, Figure 4
mosaic(MU_Direct ~ MU_BirthOrder, data=MUsub, set_labels=list(MU_Direct = c("No", "Yes"))) #direct

mosaic(MU_Indirect_dum ~ MU_BirthOrder, data=MUsub, set_labels=list(MU_Indirect_dum = c("No", "Yes"))) #indirect

```